### 2019-11-28 paper/conclusion.tex

\section{Conclusion} \label{Conclusion}

\section{Future work}

>>>>

\section{Conclusion} \label{Conclusion}
The full-core of the Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) is loaded by three different types of initial fissile materials: $^{233}$U, reactor-grade Pu, and transuranic (TRU) elements from \gls{LWR} \gls{SNF}. The excess reactivity $\rho$$_e$ is calculated at zero burnup when all CRs are fully withdrawn. The excess reactivities for $^{233}$U, reactor-grade Pu, and TRU are $1.65\pm0.04$ $\$$, $4.11\pm0.02$ $\$$, and $15.38\pm0.04$ $\$$, respectively.

Five different absorber materials are considered based on their neutronics and safety performance: B$_4$C-90, HfB$_2$, HfH$_{1.62}$, Eu$_2$O$_3$, and Gd$_2$O$_3$. Enriched B$_4$C-90 has the largest absorption ability, while Gd$_2$O$_3$ has the lowest absorption compared with the other absorber materials in this study. Both CSD and SSD clusters are separately able to shut down the reactor initially loaded by $^{233}$U and reactor-grade Pu regardless of the absorber material type. However, only SSD clusters made of B$_4$C-90 is able to shut down the SD-TMSR initially loaded by transuranic (TRU) elements.

In case of malfunction of the other CR clusters (e.g. stuck in the upper position), the outer ring of the CSD failures to counteract the excess reactivity of the core initially loaded by reactor-grade Pu and transuranic (TRU) elements. However, the worth of the outer ring of the CSD is sufficient to compensate the excess reactivity for the core refueled by $^{233}$U.

All absorber materials provide an adequate shutdown margin for the SD-TMSR core that initially loaded by $^{233}$U and reactor-grade Pu. However, the shutdown margins for TRU case are negative or slightly positive (in B$_4$C-90 case), this makes the SSD clusters ineffective to shut down the reactor loaded by TRU.

The amplification factor (A$_{CRi}$) results show that the CSD, SSD, CSD inner ring, and SSD1 are slightly amplified due to the anti-shadowing effects. The strongest anti-shadowing effect has observed in SSD4 and CSD9 clusters that are located at the boundary between the core zones with different moderator-to-fuel ratio.

The insertion of CRs deforms the radial flux shape in certain positions, i.e., around CRs positions. This shifts the neutron flux from the core center towards the periphery.

The integral and differential control rod worth are calculated for three different systems: all control rods, CSD, and SSD systems. The results show that all absorber materials have almost the same integral rod worth in the upper quarter of the core ($x$$<$$130$). Further insertion of the control rods shows the unique absorption characteristics of each material.

\section{Future work}
The effect of burnup on the absorption ability of control rods will be investigated specifically in the second part of this paper. Additionally, .......

### 2019-11-28 paper/introduction.tex

The \gls{GIF} \cite{doe2002technology} determined six innovative reactor systems: the Very High-Temperature Reactor (VHTR), the \gls{MSR}, the Supercritical Water-Cooled Reactor (SWCR), the Gas-cooled Fast Reactor (GFR), the Sodium-cooled Fast Reactor (SFR), and the Lead-cooled Fast Reactor (LFR). The MSR is the only liquid-fueled reactor among these reactors. Major nuclear centers pursue MSRs with renewed interest \cite{betzler_impacts_2019,ashraf2019whole_core,betzler2016modeling,mohsin2019safety,zhang2020radiotoxicity}. The MSR is a promising technology because of its dynamics, especially for thorium fuel utilization. However, this technology has challenges in safety, online reprocessing, and fuel handling. Safety is one of the technological goals of the MSR systems that require further research and investigation. The unique characteristics of the MSRs (liquid fuel, flux level, neutron economy, etc.) strongly affect its control system design. The MSRs have a negative temperature coefficient of reactivity. Doppler effect and thermal expansion of the fuel reduce core reactivity when the core heats up. However, graphite thermal expansion positively affect the core reactivity \cite{mathieu2006thorium}. The graphite expansion is not a dominating factor, thus core reactivity will decrease automatically when the core heats up \cite{li_optimization_2018}. The MSR designs have a drain tank to collect the liquid fuel in an emergency case. A freeze plug, which located under the core, melts when the temperature reaches a critical point. Then the fuel moves to the drain tank to shut down the reactor \cite{elsheikh2013safety}

Safety aspects in the MSRs can be listed as follows \cite{elsheikh2013safety,merle2008optimization}:
\begin{enumerate}
\item The negative coefficient of reactivity - high core temperature will reduce reactivity.
\item The fuel feed rate - decreasing the refueling rate slows down and eventually stops the nuclear chain reaction.
\item The drain tanks, which have subcritical geometry and passively cooled.
\item The control rods compensate the excess reactivity, temperature regulation, and/or shut down the reactor.
\item The continuous removing of fission products and radioactive gasses minimize the chance of radiation exposure.
\item The reactor system chamber.
\item The outer containment vessel.
\end{enumerate}

At the Beginning of Life (BOL), we must load the MSR with larger amount of fuel than that required to achieve criticality (necessary for long term core operation). This leads to excess reactivity at the BOL \cite{duderstadt650nuclear}. Additionally, during burnup, the online reprocessing and refueling leads to change the reactivity of the MSRs \cite{ashraf2019whole_core}. The most common procedure for reactor control is to insert or withdraw control rods with large neutron absorption cross section (e.g. boron and cadmium) \cite{duderstadt650nuclear}. The control rods introduce an amount of negative reactivity into the core. This negative reactivity helps to compensate the excess reactivity and adjust the power level of the core or even shut down the reactor in an emergency situation \cite{glasstone1967nuclear}. Therefore, we should estimate the reactivity worth of the control rods \cite{varvayanni2009estimation,fadaei2009control,aoyama2007core,bretscher1997computing}. The reactivity worth of control rods correlates with the interference (shadowing effects) between control rod clusters. \cite{girardin2008development,vcerba2017optimization}. 
The control rod worth and its efficiency to absorb excess positive reactivity is a subject of major interest since it directly affects reactor safety \cite{liu2018criticality,atkinson2019small,vcerba2017optimization,do2019criticality,guo2019advanced,varvayanni2009estimation}.

@@

Guo (2019), Gosset (2017), and Rudy (2011) summarized the properties of the potential alternative absorbers in Generation-IV reactors such as hafnium-based materials and rare earth element oxides. These absorbers have high thermal conductivity, good resistance to neutron irradiation, and release no gas \cite{guo2019optimized,gosset2017absorber,konings2011comprehensive}.

The \gls{SD-TMSR} with a thermal power of 2250 MW$_{th}$ \cite{ashraf2019whole_core,li_optimization_2018} is a graphite-moderated molten salt reactor. Recent studies on MSRs show that the excess reactivity at the beginning of the operation is quite large for many fueling scenarios \cite{ashraf2019whole_core,rykhlevskii2019modeling,betzler2016modeling,ashraf2018nuclear,ashraf2019modeling}. Adjusting fertile and fissile feed rates helps to control the reactivity of the SD-TMSR \cite{ashraf2019whole_core}. However, we should introduce a reliable safety system beside the online feed system to control the excess reactivity. Therefore, the main objective of our study is to introduce a new safety system based on control rods in the \gls{SD-TMSR} to control the reactivity during normal operation. Five different absorber materials are applied to study the main neutronics and safety parameters in the SD-TMSR. We focus on the control rod design, absorption ability, integral and differential control rod worths, shutdown margin of absorber materials, and shadowing effects at steady state calculation.

>>>>

The \gls{GIF} \cite{doe2002technology} determined six innovative reactor systems: the Very High-Temperature Reactor (VHTR), the \gls{MSR}, the Supercritical Water-Cooled Reactor (SWCR), the Gas-cooled Fast Reactor (GFR), the Sodium-cooled Fast Reactor (SFR), and the Lead-cooled Fast Reactor (LFR). The MSR is the only liquid-fueled reactor among these reactors. Major nuclear centers pursue MSRs with renewed interest \cite{betzler_impacts_2019,ashraf2019whole_core,betzler2016modeling,mohsin2019safety,zhang2020radiotoxicity}. The MSR is a promising technology because of its dynamics, especially for thorium fuel utilization. However, this technology has challenges in safety, online reprocessing, and fuel handling. Safety is one of the technological goals of the MSR systems that require further research and investigation. The unique characteristics of the MSRs (liquid fuel, flux level, neutron economy, etc.) strongly affect its control system design. Graphite-moderated MSRs have a negative temperature coefficient of reactivity \cite{robertson_conceptual_1971,nuttin2005potential,rykhlevskii2019modeling,ashraf2019whole_core}. Doppler effect and thermal expansion of the fuel reduce core reactivity when the core heats up. However, graphite thermal expansion positively affect the core reactivity \cite{mathieu2006thorium}. The graphite expansion is not a dominating factor, thus core reactivity will decrease automatically when the core heats up \cite{li_optimization_2018}. MSR designs have a drain tank(s) to collect the liquid fuel in an emergency case. A freeze plug, which located under the core, melts when the temperature reaches a critical point. Then the fuel moves to the drain tank to shut down the reactor \cite{elsheikh2013safety}.

Safety aspects in the MSRs can be listed as follows \cite{elsheikh2013safety,merle2008optimization}:
\begin{enumerate}
\item The negative coefficient of reactivity - high core temperature will reduce reactivity.
\item The fuel feed rate - decreasing the refueling rate slows down and eventually stops the nuclear chain reaction.
\item The drain tanks, which have subcritical geometry and passively cooled.
\item The control rods compensate the excess reactivity, regulate temperature, and/or shut down the reactor.
\item The continuous removing of fission products and radioactive gasses minimize the chance of radiation exposure.
\item The outer containment vessel.
\end{enumerate}

At the Beginning of Life (BOL), we must load the MSR with larger amount of fuel than that required to achieve criticality (necessary for long term core operation). This leads to excess reactivity at the BOL \cite{duderstadt650nuclear}. Additionally, during burnup, the online reprocessing and refueling leads to change the reactivity of the MSRs \cite{ashraf2019whole_core}. The most common procedure for reactor control is to insert or withdraw control rods with large neutron absorption cross section (e.g. boron and cadmium) \cite{duderstadt650nuclear}. Insertion of control rods introduces an amount of negative reactivity into the core. This negative reactivity helps to compensate the excess reactivity and adjust the power level of the core or even shut down the reactor in an emergency situation \cite{glasstone1967nuclear}. Therefore, we should estimate the reactivity worth of the control rods \cite{varvayanni2009estimation,fadaei2009control,aoyama2007core,bretscher1997computing}. The reactivity worth of control rods correlates with the interference (shadowing effects) between control rod clusters. \cite{girardin2008development,vcerba2017optimization}. 
The control rod worth and its efficiency to absorb excess positive reactivity is a subject of major interest since it directly affects reactor safety \cite{liu2018criticality,atkinson2019small,vcerba2017optimization,do2019criticality,guo2019advanced,varvayanni2009estimation}.

@@

Guo (2019), Gosset (2017), and Rudy (2011) summarized the properties of the potential alternative absorbers in Generation-IV reactors such as hafnium-based materials and rare earth element oxides. These absorbers have high thermal conductivity, good resistance to neutron irradiation, and release almost no gas \cite{guo2019optimized,gosset2017absorber,konings2011comprehensive}.

The \gls{SD-TMSR} with a thermal power of 2,250 MW$_{th}$ \cite{ashraf2019whole_core,li_optimization_2018} is a graphite-moderated molten salt reactor. Recent studies on MSRs show that the excess reactivity at the beginning of the operation is quite large for many refueling scenarios \cite{ashraf2019whole_core,rykhlevskii2019modeling,betzler2016modeling,ashraf2018nuclear,ashraf2019modeling}. Adjusting fertile and fissile feed rates helps to control the reactivity of the SD-TMSR \cite{ashraf2019whole_core}. However, we should introduce a reliable safety system beside the online feed system to control the excess reactivity. Therefore, the main objective of our study is to introduce a new safety system based on control rods in the \gls{SD-TMSR} to control the reactivity during normal operation. Five different absorber materials are applied to study the main neutronics and safety parameters in the SD-TMSR. We focus on the control rod design, absorption ability, integral and differential control rod worths, shutdown margin of absorber materials, and shadowing effects at steady state calculation.

### 2019-11-28 paper/results.tex

The full-core of the Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) is loaded by three different types of initial fissile materials: $^{233}$U, reactor-grade Pu, and transuranic (TRU) elements from \gls{LWR} \gls{SNF}.
The excess reactivity $\rho$$_e$ is calculated at zero burnup (steady state calculation), when all CRs are fully withdrawn. The excess reactivities for $^{233}$U, reactor-grade Pu, and TRU are $1.65\pm0.04$ $\$$, $4.11\pm0.02$ $\$$, and $15.38\pm0.04$ $\$$, respectively. For $^{233}$U case, the maximum excess reactivity $\rho$$_e$ is about $4.27\pm0.01$ $\$$ during 60 effective full-power years (EFPY) of reactor operation (see Figure~\ref{fig:keff_25}). The SD-TMSR is able to control the reactivity by adjusting the online refueling and reprocessing rates. An effective and reliable reactivity control system must be designed beside the online feed system to operate with such reactivity changes during burnup.

@@

The total worth of all control rods ranges from $33.7\pm0.4$ to $48.19\pm0.68$ \$ (Table~\ref{tab:worth}). Enriched B$_4$C-90 has the largest absorption ability, while Gd$_2$O$_3$ has the lowest absorption compared with the other absorber materials in this study. This result agrees with macroscopic absorption cross sections data \cite{guo2019optimized}. Enriched B$_4$C-90 has the highest macroscopic absorption cross sections followed by Eu$_2$O$_3$, HfB$_2$, HfH$_{1.62}$, and finally Gd$_2$O$_3$. The absorber materials are consumed during burnup. The effect of burnup on absorption ability will be investigated specifically in the second part of this paper.

@@

In case of malfunction of the other CR clusters (e.g. stuck in the upper position), the outer ring of the CSD failures to counteract the excess reactivity of the core initially loaded by reactor-grade Pu and transuranic (TRU) elements. However, the worth of the outer ring of the CSD is sufficient to compensate the excess reactivity for the core with $^{233}$U refueled.

@@

The SSD clusters are designed mainly for an emergency shutdown, thus it should provide the reactor core with sufficient and adequate negative reactivity. The shutdown margin (SDM) is calculated by equation~\ref{Equ:6}. Table~\ref{tab:table2} summarizes the shutdown margins for the SD-TMSR core initially loaded by $^{233}$U,  reactor-grade Pu, and transuranic (TRU) elements for different absorber materials.
All absorber materials provide an adequate shutdown margin for the SD-TMSR core that initially loaded by $^{233}$U and reactor-grade Pu. However, the shutdown margins for TRU case are negative or slightly positive (in B$_4$C-90 case), this makes the SSD clusters ineffective to shut down the reactor in these cases.

@@

Figure~\ref{fig:CSD} shows the integral CRW for only CSD clusters with five different absorber materials. The results show that all absorber materials have almost the same integral rod worth in the upper quarter of the core ($x$<$130$). Further insertion of the control rod clusters shows the unique absorption characteristics of each material see part \ref{CR_worth}. Enriched B$_4$C-90 absorbs neutrons much grater than the Gd$_2$O$_3$ that has the lowest absorption ability among the other absorber materials in this study. All result are based on steady state calculations. Further and detailed analysis including burnup calculation will be represented in the near future.

>>>>

The full-core of the Single-fluid Double-zone Thorium-based Molten Salt Reactor (SD-TMSR) is loaded by three different types of initial fissile materials: $^{233}$U, reactor-grade Pu, and transuranic (TRU) elements from \gls{LWR} \gls{SNF}.
The excess reactivity $\rho$$_e$ is calculated at zero burnup (steady state calculation), when all CRs are fully withdrawn. The excess reactivities for $^{233}$U, reactor-grade Pu, and TRU are $1.65\pm0.04$ $\$$, $4.11\pm0.02$ $\$$, and $15.38\pm0.04$ $\$$, respectively. For $^{233}$U case, the maximum excess reactivity $\rho$$_e$ is about $4.27\pm0.01$ $\$$ during 60 effective full-power years (EFPY) of reactor operation (see Figure~\ref{fig:keff_25}). The SD-TMSR is able to control the reactivity by adjusting the online refueling and reprocessing rates \cite{ashraf2019whole_core}. An effective and reliable reactivity control system must be designed beside the online feed system to operate with such reactivity changes during burnup.

@@

The total worth of all control rods ranges from $33.7\pm0.4$ to $48.19\pm0.68$ \$ (Table~\ref{tab:worth}). Enriched B$_4$C-90 has the largest absorption ability, while Gd$_2$O$_3$ has the lowest absorption compared with the other absorber materials in this study. This result agrees with macroscopic absorption cross sections data \cite{guo2019optimized}. Enriched B$_4$C-90 has the highest macroscopic absorption cross sections followed by Eu$_2$O$_3$, HfB$_2$, HfH$_{1.62}$, and finally Gd$_2$O$_3$. The absorber materials are consumed during burnup. The effect of burnup on the absorption ability will be investigated specifically in the second part of this paper.

@@

In case of malfunction of the other CR clusters (e.g. stuck in the upper position), the outer ring of the CSD failures to counteract the excess reactivity of the core initially loaded by reactor-grade Pu and transuranic (TRU) elements. However, the worth of the outer ring of the CSD is sufficient to compensate the excess reactivity for the core refueled by $^{233}$U .

@@

The SSD clusters are designed mainly for an emergency shutdown, thus it should provide the reactor core with sufficient and adequate negative reactivity. The shutdown margin (SDM) is calculated by equation~\ref{Equ:6}. Table~\ref{tab:table2} summarizes the shutdown margins for the SD-TMSR core initially loaded by $^{233}$U,  reactor-grade Pu, and transuranic (TRU) elements for different absorber materials.
All absorber materials provide an adequate shutdown margin for the SD-TMSR core that initially loaded by $^{233}$U and reactor-grade Pu. However, the shutdown margins for TRU case are negative or slightly positive (in B$_4$C-90 case), this makes the SSD clusters ineffective to shut down the reactor loaded by TRU.

@@

Figure~\ref{fig:CSD} shows the integral CRW for only CSD clusters with five different absorber materials. The results show that all absorber materials have almost the same integral rod worth in the upper quarter of the core ($x$$<$$130$). Further insertion of the control rods shows the unique absorption characteristics of each material see part \ref{CR_worth}. Enriched B$_4$C-90 absorbs neutrons much grater than the Gd$_2$O$_3$ that has the lowest absorption ability among the other absorber materials in this study. All result are based on steady state calculations. Further and detailed analysis including burnup calculation will be represented in the near future.
